Getting Started with the Graded Response Model (GRM): An introduction and tutorial in R

### 2024-10-24 annotated_codes.Rmd
final revision

---
title: "Tutorial on running a Graded Response Model"
author: "Rizqy Amelia Zein & Muhamad Hanif Akhtar"
date: "`r Sys.Date()`"
output: html_document
---

@@

The simulated data shows the same pattern as the real one, which shows that the scale as a whole is optimal (having sufficient reliability, \~0.8) for measuring participants with theta levels between -2SD and +4SD.

@@

If you have any feedback, please get in touch and let us know by email here: [amelia.zein\@psikologi.unair.ac.id](mailto:amelia.zein@psikologi.unair.ac.id).

>>>>

---
title: "Tutorial on running a Graded Response Model"
author: "Rizqy Amelia Zein & Hanif Akhtar"
date: "`r Sys.Date()`"
output: html_document
---

@@

The simulated data shows the same pattern as the real one, which shows that the scale as a whole is optimal (having sufficient reliability, \~0.8) for measuring participants with $\theta$ levels between -2SD and +4SD.

## Contact

@@

If you have any feedback or questions, please get in touch and let us know by email: [Amelia.Zein\@psy.lmu.de](mailto:Amelia.Zein@psy.lmu.de?subject=GRM%20Tutorial), [amelia.zein\@psikologi.unair.ac.id](mailto:amelia.zein@psikologi.unair.ac.id?subject=GRM%20Tutorial), or [hanifakhtar\@umm.ac.id](mailto:hanifakhtar@umm.ac.id?subject=GRM%20Tutorial).

### 2024-10-24 manuscript.qmd
final revision

---
title: "Getting Started with the Graded Response Model (GRM): An introduction and tutorial in R"
shorttitle: "The Graded Response Model Tutorial"
author:
  - name: Rizqy Amelia Zein
    corresponding: true
    orcid: 0000-0001-7840-0299
    email: Amelia.Zein@psy.lmu.de
    affiliations:
      - name: Ludwig-Maximilians-Universität
        department: Department of Psychology
        address: Leopoldstraße 13
        city: Munich
        region: Germany
        postal-code: "80802"
      - name: Universitas Airlangga
        department: Department of Psychology
        address: Jalan Airlangga 4-6
        city: Surabaya, Indonesia
        postal-code: "60286"
  - name: Hanif Akhtar
    orcid: 0000-0002-1388-7347
    affiliations:
      - name: Universitas Muhammadiyah Malang
        department: Faculty of Psychology
        city: Malang
        region: Jawa Timur, Indonesia
abstract: "This tutorial introduces the Graded Response Model (GRM), a tool for testing measurement precision within the Item Response Theory (IRT) paradigm, which is useful for informing researchers about the item and person properties of their measurement. The tutorial aims to guide applied researchers through a unidimensional GRM analysis in the R environment, using the psych, mirt, and ggmirt packages. GRM is specifically designed to examine the psychometric properties of psychological scales with polytomous (Likert-style) items. The tutorial illustrates the procedure using data from the Open Psychometrics Database on the Right-Wing Authoritarianism (RWA) scale, outlining the theoretical underpinnings of GRM and steps for data preparation, testing model assumptions, model fitting, plotting item parameters, and interpretation of results."
keywords:
  - graded response model
  - item response theory
  - mirt
  - ggmirt
  - psych
  - R
author-note:
  disclosures:
    conflict-of-interest: "We declare that we have no competing interests."
    financial-support: "The first author receives a PhD scholarship from the German Academic Exchange Service (Deutscher Akademischer Austauschdienst, DAAD), award number 91803023."
lang: en
bibliography: references.bib
format:
  apaquarto-docx: default
editor: visual
execute: 
  warning: false
  error: false
  output: false
  echo: true
---

@@

| Model                       | Key Characteristics                                                                                                                                                                                                                                          | Data Type   | Response Options        |
|------------|-------------------------------------|------------|------------|
| 1-PL Model (Rasch Model)    | 1\. Estimates only item difficulties (*b*). 2. Assumes that all items have equal discrimination parameters (*a*). 3. Item and person parameters are independent.                                                                                             | Dichotomous | Correct/Incorrect (0/1) |
| 2-PL Model                  | 1\. Estimates item difficulties (*b*) and item discrimination (*a*). 2. Less stringent than 1-PL model since it allows item discrimination parameters (*a*) to vary.                                                                                         | Dichotomous | Correct/Incorrect (0/1) |
| 3-PL Model                  | 1\. Estimates item difficulties (*b*), discrimination (*a*), and pseudo-guessing parameter (*c*). 2. Appropriate for modeling a test data with multiple responses (e.g., multiple-choice tests), and thus, guessing might influence participants' responses. | Dichotomous | Correct/Incorrect (0/1) |
| Graded Response Model (GRM) | 1\. Appropriate for modeling ordinal data with more than two response categories (i.e., Likert-style). 2. Estimates a discrimination parameter (*a*) and multiple threshold parameters (*b*) per item.                                                       | Polytomous  | Ordered Categories      |

@@

To maximize the reproducibility of our illustrative example, we wrote the article as a Quarto (.qmd) document, where we integrate the R codes used in the analysis as well as its outputs. We also include a longer, annotated R Markdown file (.Rmd) for the example we used as a supplementary document, which we highly recommend using for a didactic purpose. In the same file, we also include a part where we demonstrate the application of GRM using a simulated dataset, which we do not include in the article, so that readers can compare the example we present here against an ideal scenario. The Quarto file (and its corresponding .docx and .pdf output) and an annotated R Markdown file are publicly available on [a Github repository]() (*unlinked for blind review*). The dataset we used in this tutorial paper was obtained from a publicly accessible [Open-Source Psychometrics Project](http://openpsychometrics.org/_rawdata/).

@@

Before running the analysis, we need to install the necessary R packages, which are, *mirt* [@chalmersMirtMultidimensionalItem2023a; @chalmersMirtMultidimensionalItem2012], *psych* [@revellePsychProceduresPsychological2023], *ggmirt* [@masurGgmirt2023], and *caret* [@kuhnCaretClassificationRegression2024], and *skimr* [@waringSkimrCompactFlexible2024] with these following commands:

>>>>

---
title: "Getting Started with the Graded Response Model (GRM): An introduction and tutorial in R"
shorttitle: "The Graded Response Model Tutorial"
author:
  - name: Rizqy Amelia Zein
    corresponding: true
    orcid: 0000-0001-7840-0299
    email: Amelia.Zein@psy.lmu.de
    affiliations:
      - name: Ludwig-Maximilians-Universität
        department: Department of Psychology
        address: Leopoldstraße 13
        city: Munich
        region: Germany
        postal-code: "80802"
      - name: Universitas Airlangga
        department: Department of Psychology
        address: Jalan Airlangga 4-6
        city: Surabaya, Indonesia
        postal-code: "60286"
  - name: Hanif Akhtar
    orcid: 0000-0002-1388-7347
    affiliations:
      - name: Universitas Muhammadiyah Malang
        department: Faculty of Psychology
        city: Malang
        region: Jawa Timur, Indonesia
abstract: "This tutorial introduces the Graded Response Model (GRM), a tool for testing measurement precision within the Item Response Theory (IRT) paradigm, which is useful for informing researchers about the item and person properties of their measurement. The tutorial aims to guide applied researchers through a unidimensional GRM analysis in the R environment, using the psych, mirt, and ggmirt packages. GRM is specifically designed to examine the psychometric properties of psychological scales with polytomous (Likert-style) items. The tutorial illustrates the procedure using data from the Open Psychometrics Database on the Right-Wing Authoritarianism (RWA) scale, outlining the theoretical underpinnings of GRM and steps for data preparation, testing model assumptions, model fitting, plotting item parameters, and interpretation of results."
keywords:
  - graded response model
  - item response theory
  - mirt
  - ggmirt
  - psych
  - R
author-note:
  disclosures:
    conflict-of-interest: "We declare that we have no competing interests."
    financial-support: "The first author receives a PhD scholarship from the German Academic Exchange Service (Deutscher Akademischer Austauschdienst, DAAD), award number 91803023, and thanks DAAD for its financial support."
lang: en
bibliography: references.bib
format:
  apaquarto-docx: default
editor: visual
execute: 
  warning: false
  error: false
  output: false
  echo: true
---

@@

| Model                       | Key Characteristics                                                                                                                                                                                                                                          | Data Type   | Response Options        |
|-----------------------------|--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|-------------|-------------------------|
| 1-PL Model (Rasch Model)    | 1\. Estimates only item difficulties (*b*). 2. Assumes that all items have equal discrimination parameters (*a*). 3. Item and person parameters are independent.                                                                                             | Dichotomous | Correct/Incorrect (0/1) |
| 2-PL Model                  | 1\. Estimates item difficulties (*b*) and item discrimination (*a*). 2. Less stringent than 1-PL model since it allows item discrimination parameters (*a*) to vary.                                                                                         | Dichotomous | Correct/Incorrect (0/1) |
| 3-PL Model                  | 1\. Estimates item difficulties (*b*), discrimination (*a*), and pseudo-guessing parameter (*c*). 2. Appropriate for modeling a test data with multiple responses (e.g., multiple-choice tests), and thus, guessing might influence participants' responses. | Dichotomous | Correct/Incorrect (0/1) |
| Graded Response Model (GRM) | 1\. Appropriate for modeling ordinal data with more than two response categories (i.e., Likert-style). 2. Estimates a discrimination parameter (*a*) and multiple threshold parameters (*b*) per item.                                                       | Polytomous  | Ordered Categories      |

@@

To maximize the reproducibility of our illustrative example, we wrote the article as a Quarto (.qmd) document [@schneiderWjschneApaquarto2024], where we integrate the R codes used in the analysis as well as its outputs. We also include a longer, annotated R Markdown file (.Rmd) for the example we used as a supplementary document, which we highly recommend using for a didactic purpose. In the same file, we also include a part where we demonstrate the application of GRM using a simulated dataset, which we do not include in the article, so that readers can compare the example we present here against an ideal scenario. The Quarto file (and its corresponding .docx and .pdf output) and an annotated R Markdown file are publicly available on [a Github repository](https://github.com/rameliaz/grm-tutorial-paper). The dataset we used in this tutorial paper was obtained from a publicly accessible [Open-Source Psychometrics Project](http://openpsychometrics.org/_rawdata/).

@@

Before running the analysis, we need to install the necessary R packages, which are, *mirt* [@chalmersMirtMultidimensionalItem2012], *psych* [@revellePsychProceduresPsychological2023], *ggmirt* [@masurGgmirt2023], and *caret* [@kuhnCaretClassificationRegression2024], and *skimr* [@waringSkimrCompactFlexible2024] with these following commands:
